---
title: "Project title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Data 1

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

Based on agriculture, industry, and services, how has the Gross Value Added changed for Central and Latin American countries between 1995 and 2020? 

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

Since our dataset states the gross value added (GVA) of all countries, we wanted to focus on a particular region within the world, in this case, Central and Latin America, particularly because they had changes to their economies due to the Latin American Boom in the twentieth century. By analyzing the GVA changes throughout time, we will be able to analyze how countries\' economies have changed and what this says about the wealth of the economy at the time. Our hypothesis is that due to the Latin American Boom, most Central and Latin American countries\' GVA generally increased over time.

-   Identify the types of variables in your research question. Categorical? Quantitative?

The types of variables in our research question are categorical ordinal because the columns are dates. 

## Literature

-   Find one published credible article on the topic you are interested in researching.

Brenes, Esteban R., Arnoldo R. Camacho, Luciano Ciravegna, and Caleb A. Pichardo. \"Strategy and Innovation in Emerging Economies after the End of the Commodity Boom---Insights from Latin America.\" Journal of Business Research 69, no. 10 (February 1, 2016): 4363--67. https://doi.org/10.1016/j.jbusres.2016.03.059. 

-   Provide a one paragraph summary about the article.

The article explores the economic trends in Latin America, highlighting their influences on the strategies and behaviors of different types of businesses. The article offers contextual information about Latin America\'s positive economic performance between the 1990s and 2010s, which aligns perfectly with the UN data set. Moreover, the article offers insights into Latin America\'s emerging markets throughout this time period. The article utilizes information from the Sixth International Conference on Strategic Management in Latin America, providing an opportunity to exchange empirical research results about the strategic economy.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

Our research question builds on the article we have cited because it provides direct empirical evidence relating to the context offered by the conference in Latin America (discussed by the article).

## Glimpse of data

```{r}
#| label: load-data-1

gva_un <- read.csv("data/GVA_Altered.csv")
```

# Data 2

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-2

land_un <- read.csv("data/Land_UN_Altered.csv")
```

# Data 3

## Introduction and data

-   Identify the source of the data.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

-   Write a brief description of the observations.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

-   A description of the research topic along with a concise statement of your hypotheses on this topic.

-   Identify the types of variables in your research question. Categorical? Quantitative?

## Literature

-   Find one published credible article on the topic you are interested in researching.

-   Provide a one paragraph summary about the article.

-   In 1-2 sentences, explain how your research question builds on / is different than the article you have cited.

## Glimpse of data

```{r}
#| label: load-data-3

homicides_and_crime_un <- read.csv("data/Homicides_Other_Crimes_UN_Altered.csv")
```
