---
title: "International Crime Rates"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

# **Research Project: International Crime**

# **Introduction:**

Many consider the 21st century as the best time to be human as we are afforded luxuries and lifestyles that have never before been so accessible. But is this world safer? As technology and society develop, income inequality has skyrocketed (Kemal & Qureshi 2016). The media and news cycle show their viewers a world that is increasingly more dangerous, less safe, and rife with criminal activity. Our research seeks to answer the following research questions: 

# **Research Question(s):**

**Has the world become more violent in the past 15 years, and if so, how does this breakdown across region? Given this, does GDP correlate with these observations?**

Our **hypothesis** is that contrary to portrayals in media, much of the world is actually becoming less violent. We **also hypothesize** that volatile and unstable regions of the world, whether that be from civil conflict, interstate conflict, eroding rule of law, or political strife, are responsible for a large share of the violent crimes and homicides that fold into the worldwide statistics. This research will also allow us to **empirically evaluate** a 1976 UN Social Defense Research Institute claim regarding the often-cited observation that GDP and violent crime are correlated, if only at the macro level (Luna et al., 1976). Given this claim, we believe that as a region's GDP increases, their rates of violent crimes decrease.  

The **homicide and crime data** was originally curated by the United Nations Office on Drugs and Crime (ODC) Homicide Statistics. The ODC compiles national and international data on homicide from a variety of sources (including criminal justice and public health systems from the countries included and the World Health Organization) which include data that conform to the ODC's definition of crime, that being '' the punishable contravention or violation of the limits on human behavior as imposed by national criminal legislation." Homicides are defined as the intentional killing of one person by another by the UN. The data is recorded in 5-year intervals with the first publication being in 2005 and the most recent being 2020. There are ethical concerns resulting from the fact that every one of those millions that make up the statistics of deaths and crime are still personal tragedies. Therefore, a certain amount of sensitivity around the subject is required. Also, many countries may not be as transparent with their homicide statistics, whether for political reasons or poor infrastructure around research, which could reduce the accuracy of the results. Therefore, this study acknowledges the possible weaknesses in data collection.

The **GDP data** was originally compiled by the UN Statistics Division and shows the GDP of each country through the years 1995-2020 ( in approx. 5 year intervals), demonstrating how many countries have managed to increase their GDP substantially. Since this data set is merely factual representations of the economic growth of the given nations, it did not pose any discernible threat to the individuals in these countries. Since no civilian was impacted by the methods used to obtain the data, there are no ethical concerns.

# **Data Insertion + Transformation for Analysis:**

```{r}
# | label: data-inseration

library(tidyverse)
library(tidymodels)

replaceCommas <- function(x)
  {x <- as.numeric(gsub("\\,", "", x))} #cited

gdp_un <- read.csv("data/GDP_GDPpC_UN_Altered.csv", na.strings = ",")
gdp_un <- gdp_un |>
  rename("SeriesGDP" = "Series", "ValueGDP" = "Value", "FootnotesGDP" = "Footnotes", "XGDP" = "X", "Region.Country.Area.GDP" = "Region.Country.Area")
gdp_un$ValueGDP <- replaceCommas(gdp_un$ValueGDP)
gdp_un$ValueGDP <- as.numeric(as.character(gdp_un$ValueGDP))

homicides_and_crime_un <- read.csv("data/Homicides_Other_Crimes_UN_Altered.csv")
homicides_and_crime_un$Value <- as.numeric(as.character(homicides_and_crime_un$Value))
```

**Citation:** https://stackoverflow.com/questions/49910861/removing-comma-from-numbers-in-r

# **Methodology:**

For the data analysis, we intend to utilize visualizations of the regional and global homicide statistics to derive our findings as related to our first question. For the latter half of our question, we will be utilizing a series of regression models to determine the level of correlation between homicides and GDP via their R\^2 values.

#### All Regions, Combined:

```{r}
#| label: regions-combined

Total_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Oceania" |
         X == "Africa" |
         X == "Europe" |
         X == "Asia" |
         X == "Latin America & the Caribbean" |
         X == "Northern America") |>
  filter(Series == "Intentional homicide rates per 100,000") |>
  mutate(Value = as.numeric(Value), X = as.factor(X))

Total_homicides_and_crime_un |>
  ggplot(aes(
    x = Year, 
    y = Value,
    color = X)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(
    breaks = c(0,5,10,15,20,25), labels = c(0,5,10,15,20,25)
    ) +
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Regions, Combined",
    color = "Region") +
  theme_bw() +
  scale_color_viridis_d()
```

**Analysis:**

We chose to include this graph with all these countries in one visual representation, so it is easier to see the different patterns among countries, as there is a consistent scale of reference. For example, homicide victims have decreased in Latin America & the Caribbean but the overall amount of homicide victims is significantly higher than all other regions. Also, near the bottom of the graph, there is a cluster of regions, including the Northern America, Oceania, Asia, and Europe. This cluster demonstartes that similar crime patterns in different regions of the world. Futhermore, Africa is seen in the middle of graph, with its homicide rate remaining pretty constant compared to all other regions. However, more detailed patterns are necessary to reject or support our hypothesis. Therefore, the study includes the specific patterns of each country later on.

### **R.3 - Global - R\^2**

```{r}
# | label: total-gdp-homicides

gdp_un_total <- gdp_un |>
  filter(Region.Country.Area.GDP == "1") |>
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)")
gdp_un_total$ValueGDP <- as.numeric(as.character(gdp_un_total$ValueGDP))

#Crime
homicides_and_crime_un_total <- homicides_and_crime_un |>
  filter(Region.Country.Area == "1") |>
  filter(Series == "Intentional homicide rates per 100,000") 

#GDP
gdp_current_and_crime_total <- homicides_and_crime_un_total |> 
  full_join(gdp_un_total, by = "Year") |> 
  filter(!is.na(Region.Country.Area)) 

modeltotal <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_total)

glance(modeltotal)$r.squared
```

### **R.3 - Regional - R\^2**

#### Africa

```{r}
#| label: africa-regression&r^2

gdp_un_africa <- gdp_un |>
  filter(Region.Country.Area.GDP == "2") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_africa$ValueGDP <- as.numeric(as.character(gdp_un_africa$ValueGDP))

homicides_and_crime_un_africa <- homicides_and_crime_un |>
  filter(Region.Country.Area == "2") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_africa <- homicides_and_crime_un_africa |> 
full_join(gdp_un_africa, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelafrica <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_africa)

glance(modelafrica)$r.squared
```

#### North America

```{r}
#| label: northamerica-regression&r^2

gdp_un_northamerica <- gdp_un |>
  filter(Region.Country.Area.GDP == "21") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)")

gdp_un_northamerica$ValueGDP <- as.numeric(as.character(gdp_un_northamerica$ValueGDP))

homicides_and_crime_un_northamerica <- homicides_and_crime_un |>
  filter(Region.Country.Area == "21") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_northamerica <- homicides_and_crime_un_northamerica |> 
full_join(gdp_un_northamerica, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelnorthamerica <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_northamerica)

glance(modelnorthamerica)$r.squared
```

#### Latin America

```{r}
#| label: latincarribean-regression&r^2

gdp_un_latincarribean <- gdp_un |>
  filter(Region.Country.Area.GDP == "419") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_latincarribean$ValueGDP <- as.numeric(as.character(gdp_un_latincarribean$ValueGDP))

homicides_and_crime_un_latincarribean <- homicides_and_crime_un |>
  filter(Region.Country.Area == "419") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_latincarribean <- homicides_and_crime_un_latincarribean |> 
full_join(gdp_un_latincarribean, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modellatincarribean <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_latincarribean)

glance(modellatincarribean)$r.squared
```

#### Asia

```{r}
#| label: asia-regression&r^2

gdp_un_asia <- gdp_un |>
  filter(Region.Country.Area.GDP == "142") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_asia$ValueGDP <- as.numeric(as.character(gdp_un_asia$ValueGDP))

homicides_and_crime_un_asia <- homicides_and_crime_un |>
  filter(Region.Country.Area == "142") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_asia <- homicides_and_crime_un_asia |> 
full_join(gdp_un_asia, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelasia <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_asia)

glance(modelasia)$r.squared
```

#### Europe

```{r}
#| label: europe-regression&r^2

gdp_un_europe <- gdp_un |>
  filter(Region.Country.Area.GDP == "150") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_europe$ValueGDP <- as.numeric(as.character(gdp_un_europe$ValueGDP))

homicides_and_crime_un_europe <- homicides_and_crime_un |>
  filter(Region.Country.Area == "150") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_europe <- homicides_and_crime_un_europe |> 
full_join(gdp_un_europe, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modeleurope <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_europe)

glance(modeleurope)$r.squared
```

#### Oceania

```{r}
#| label: oceania-regression&r^2

gdp_un_oceania <- gdp_un |>
  filter(Region.Country.Area.GDP == "9") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_oceania$ValueGDP <- as.numeric(as.character(gdp_un_oceania$ValueGDP))

homicides_and_crime_un_oceania <- homicides_and_crime_un |>
  filter(Region.Country.Area == "9") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_oceania <- homicides_and_crime_un_oceania |> 
full_join(gdp_un_oceania, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modeloceania <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_oceania)

glance(modeloceania)$r.squared
```

### **RQ.1 - Global +Regional - Homicide Statistics 2005-2020**

#### Global

```{r}
#| label: Global

Global_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(Region.Country.Area == "1") |>
  filter(Series == "Intentional homicide rates per 100,000")

Global_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims - Global") +
  theme_bw()
```

**Analysis:** It seems that globally, there was a increasingly more steep decline in violence on a 5-year basis, supporting our hypothesis.

#### Africa

```{r}
#| label: Africa

Africa_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Africa") 

Africa_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Africa") +
  theme_bw()
```

**Analysis:** It is evident that Africa experienced a period of decline in regards to homicide in the early 2000s, only to increase beginning in 2010. However, the trend seems to have reversed and the number of homicide victims is declining, albeit at a lower rate.

#### North America

```{r}
#| label: Northern-America

NAmerica_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Northern America") |>
  filter(Series == "Intentional homicide rates per 100,000")

NAmerica_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in N. America") +
  theme_bw()
```

**Analysis:** Homicide victims in North America decreased from 2005-2010; however, since 2010, there has been an increase. From 2015-2022, homicide rates increased at a steeper rate and reached new highs.

#### **Latin America**

```{r}
#| label: LA-And-Caribbean

LACarib_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Latin America & the Caribbean") |>
  filter(Series == "Intentional homicide rates per 100,000")

LACarib_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Latin America and the Caribbean") +
  theme_bw()
```

**Analysis:** In Latin America, the number of homicide victims did increase up until 2010. Since 2010, the rate has reversed its course aand begun to decrease at a steady rate.

#### **Asia**

```{r}
#| label: Asia

Asia_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Asia") |>
  filter(Series == "Intentional homicide rates per 100,000")

Asia_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Asia") +
  theme_bw()
```

**Analysis:** While homicide victims in Asia has decreased since the first year the data was recorded, it is evident that the declines have gotten steeper as the years have gone on.

#### Europe

```{r}
#| label: Europe

Europe_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Europe") |>
  filter(Series == "Intentional homicide rates per 100,000")

Europe_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Europe") +
  theme_bw()
```

**Analysis:** Similar to Asia, Europe has also seen a decrease in their homicide rate; however, the rates of the decline have varied. From 2005 - 2010, there was a steep decline, which seemed to have levelled off from 2010-2015. From 2015 onward, the number of homicide victims seems to have decreased but at a much slower rate.

#### Oceania

```{r}
#| label: Oceania

Oceania_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Oceania") |>
  filter(Series == "Intentional homicide rates per 100,000")

Oceania_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Oceania") +
  theme_bw()
```

**Analysis:** Oceania saw a decline in the homicide rates from 2005-2010, which seemed to level off from 2010-2015, but the homicide victims have increased since 2015. However, rates overall are low so these changes aren't as drastic as they appear.

# **Results:**

Overall, our findings were the following:

**Global and Regional Statistics**

Based on our initial findings, **our hypothesis is supported** by the UN data. The global homicide rates have been consistently declining on 5-year intervals. Breaking the statistics down regionally, we find that Asia, Europe and Africa are likely responsible for this observation on account of their massive share of the global population. Surprisingly, North America saw a large spike in homicides since 2010, possible on account of the 2008 financial crisis and its impact on the middle class in the US.

**Regressions - Regional and Global**

Overall, based on our initial findings, **there is mixed correlation between GDP and homicides.** The global homicide and GDP statistics had a very high R\^2 value (0.93). Breaking the statistics down regionally, we find that Asian and European statistics have high R\^2 values, African and Latin American statistics have middling R\^2 values, and Oceanian and North American statistics have very low R\^2 values. These results **call into question** the application of the Justice Department's claim that GDP and violent crime are correlated, but we do lack the full scope of the data needed to fully evaluate the claim.

# **Works Cited:**

Derviş, Kemal, and Zia Qureshi. \"Income Distribution within Countries: Rising Inequality.\" Brookings Institution Research, The Brookings Institution, 1 Aug. 2016, [https://www.brookings.edu/research/income-distribution-within-countries-rising-inequality](https://www.brookings.edu/research/income-distribution-within-countries-rising-inequality/).

\"Gross domestic product and gross domestic product per capita\" UNdata, UN Statistics Division, Mar. 2022, <http://data.un.org/_Docs/SYB/PDFs/SYB65_230_202209_GDP%20and%20GDP%20Per%20Capita.pdf>.

\"Intentional homicides and other crimes\" UNdata, United Nations Office on Drugs and Crime, Jun. 2022, <http://data.un.org/_Docs/SYB/PDFs/SYB65_328_202209_Intentional%20homicides%20and%20other%20crimes.pdf>.

Luna, Giuliana et al. \"Economic Crisis and Crime.\" Office of the U.S. Justice Department, United Nations Social Defense Research Institute, 1 Aug. 2016, <https://www.ojp.gov/pdffiles1/Digitization/35597NCJRS.pdf>.
