---
title: "Project title"
subtitle: "Report"
format: html
editor: visual
execute:
  echo: true
---

**Research Project: International Crime** Research Question:

**Has the world become more violent in the past 15 years (2005-2020), both worldwide and regionally?**

**How does sex factor into the results?**

**How do these rates correlate with regional and national GDPs?**

**Insertion of Data:**

```{r}

library(tidyverse)
library(tidymodels)
replaceCommas<-function(x){
  x<-as.numeric(gsub("\\,", "", x))}
gdp_un <- read.csv("data/GDP_GDPpC_UN_Altered.csv", header=TRUE, na.strings = ",")
gdp_un <- gdp_un |>
  rename("SeriesGDP" = "Series", "ValueGDP" = "Value", "FootnotesGDP" = "Footnotes", "XGDP" = "X", "Region.Country.Area.GDP" = "Region.Country.Area")
gdp_un$ValueGDP <- replaceCommas(gdp_un$ValueGDP)
gdp_un$ValueGDP <- as.numeric(as.character(gdp_un$ValueGDP))

homicides_and_crime_un <- read.csv("data/Homicides_Other_Crimes_UN_Altered.csv", header=TRUE)
homicides_and_crime_un$Value <- as.numeric(as.character(homicides_and_crime_un$Value))
```

**EXAMPLE FOR GETTING DATA FOR REGRESSION**

```{r}
#setting up GDP
gdp_un_total <- gdp_un |>
  filter(Region.Country.Area.GDP == "1") |> #region of interest
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") #statistic of interest
gdp_un_total$ValueGDP <- as.numeric(as.character(gdp_un_total$ValueGDP))

#setting up Crime
homicides_and_crime_un_total <- homicides_and_crime_un |>
  filter(Region.Country.Area == "1") |> #region of interest
  filter(Series == "Intentional homicide rates per 100,000") #statistic of interest

#Joining them
gdp_current_and_crime_total <- homicides_and_crime_un_total |> 
full_join(gdp_un_total, by = "Year") |> #always by Year
  filter(!is.na(Region.Country.Area)) #dropping NAs
```

**R-Squared of Data Set (Total EXAMPLE)**

modeltotal \<- linear_reg() \|\>

set_engine("lm") \|\>

fit(ValueGDP \~ Value, data = gdp_current_and_crime_total)

glance(modeltotal)\$r.squared

```{r}

modeltotal <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_total)

glance(modeltotal)$r.squared
```

**Total Linear Regression Model (With GDP) (EXAMPLE):**

```{r}

model_total <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_total)
```

**Has the world become more violent in the past 15 years (2005-2020), both worldwide and regionally, based on sex?**

**Regional Regressions**

```{r}
#| label: africa-regression&r^2
gdp_un_africa <- gdp_un |>
  filter(Region.Country.Area.GDP == "2") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_africa$ValueGDP <- as.numeric(as.character(gdp_un_africa$ValueGDP))

homicides_and_crime_un_africa <- homicides_and_crime_un |>
  filter(Region.Country.Area == "2") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_africa <- homicides_and_crime_un_africa |> 
full_join(gdp_un_africa, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelafrica <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_africa)

glance(modelafrica)$r.squared
```

```{r}
#| label: northamerica-regression&r^2

gdp_un_northamerica <- gdp_un |>
  filter(Region.Country.Area.GDP == "21") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_northamerica$ValueGDP <- as.numeric(as.character(gdp_un_northamerica$ValueGDP))

homicides_and_crime_un_northamerica <- homicides_and_crime_un |>
  filter(Region.Country.Area == "21") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_northamerica <- homicides_and_crime_un_northamerica |> 
full_join(gdp_un_northamerica, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelnorthamerica <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_northamerica)

glance(modelnorthamerica)$r.squared

```

```{r}
#| label: latincarribean-regression&r^2

gdp_un_latincarribean <- gdp_un |>
  filter(Region.Country.Area.GDP == "419") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_latincarribean$ValueGDP <- as.numeric(as.character(gdp_un_latincarribean$ValueGDP))

homicides_and_crime_un_latincarribean <- homicides_and_crime_un |>
  filter(Region.Country.Area == "419") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_latincarribean <- homicides_and_crime_un_latincarribean |> 
full_join(gdp_un_latincarribean, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modellatincarribean <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_latincarribean)

glance(modellatincarribean)$r.squared

```

```{r}
#| label: asia-regression&r^2

gdp_un_asia <- gdp_un |>
  filter(Region.Country.Area.GDP == "142") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_asia$ValueGDP <- as.numeric(as.character(gdp_un_asia$ValueGDP))

homicides_and_crime_un_asia <- homicides_and_crime_un |>
  filter(Region.Country.Area == "142") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_asia <- homicides_and_crime_un_asia |> 
full_join(gdp_un_asia, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modelasia <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_asia)

glance(modelasia)$r.squared

```

```{r}
#| label: europe-regression&r^2

gdp_un_europe <- gdp_un |>
  filter(Region.Country.Area.GDP == "150") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_europe$ValueGDP <- as.numeric(as.character(gdp_un_europe$ValueGDP))

homicides_and_crime_un_europe <- homicides_and_crime_un |>
  filter(Region.Country.Area == "150") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_europe <- homicides_and_crime_un_europe |> 
full_join(gdp_un_europe, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modeleurope <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_europe)

glance(modeleurope)$r.squared

```

```{r}
#| label: oceania-regression&r^2

gdp_un_oceania <- gdp_un |>
  filter(Region.Country.Area.GDP == "9") |> 
  filter(SeriesGDP == "GDP in current prices (millions of US dollars)") 

gdp_un_oceania$ValueGDP <- as.numeric(as.character(gdp_un_oceania$ValueGDP))

homicides_and_crime_un_oceania <- homicides_and_crime_un |>
  filter(Region.Country.Area == "9") |> 
  filter(Series == "Intentional homicide rates per 100,000")

gdp_current_and_crime_oceania <- homicides_and_crime_un_oceania |> 
full_join(gdp_un_oceania, by = "Year") |>
  filter(!is.na(Region.Country.Area))

modeloceania <- linear_reg() |>
  set_engine("lm") |>
  fit(ValueGDP ~ Value, data = gdp_current_and_crime_oceania)

glance(modeloceania)$r.squared

```

1.  Worldwide violence, based on sex: **(FIXED, EXAMPLE)**

```{r}
#| label: worldwide-violence-sex

homicides_and_crime_un_sex <- homicides_and_crime_un |>
  filter(Region.Country.Area == "1") |> #region of interest
  mutate(sex = if_else(str_detect(Series, "Male"), "Male", "Female")) |> #not used elsewhere
  filter(Series != "Intentional homicide rates per 100,000") #statistic of interest, watch the != vs == for what you want
 

homicides_and_crime_un_sex |> #most of this straight forward, make sure to include correct titling/subtitles and caption it
  group_by(sex) |>
  ggplot(aes(x = Year, y = Value, color = sex)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "# of International Homicide Victims (per 100,000 persons)",
    title = "International Homicide Victims* by Sex",
    fill = "Sex", subtitle = "Rate per 100,000 (of given sex's population)",
    caption = "*Homicides are defined as intentional homicides by this dataset")
```

2.  Regional Violence, General:

```{r}
#| label: Africa

Africa_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Africa") 

Africa_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Africa"
  ) + theme_bw()

```

2.  It is evident that Africa experienced a period of decline in regards to homicide in the early 2000s, only to increase beginning in 2010. However, the trend seems to have revered and the number of homicide victims is declining, albeit at a lower rate.

```{r}
#| label: Northern-America

NAmerica_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Northern America") |>
  mutate(
    sex = if_else(
      str_detect(Series, "Male"), "Male", "Female")
    ) |>
  filter(Series == "Intentional homicide rates per 100,000")

NAmerica_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in N. America"
  ) + theme_bw()

```

Homicide victimis in North America decreased from 2005-2010; however, since 2010, there has been an increase. From 2015-2022, homicide rates increased at a steeper rate and reached new highs.

```{r}
#| label: LA-And-Caribbean

LACarib_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Latin America & the Caribbean") |>
  mutate(
    sex = if_else(
      str_detect(Series, "Male"), "Male", "Female")
    ) |>
  filter(Series == "Intentional homicide rates per 100,000")

LACarib_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Latin America and the Caribbean"
  ) + theme_bw()
```

In Latin America, the number of homicide victims did increase up until 2010. Since 2010, the rate has reversed its course aand begun to decrease at a steady rate.

```{r}
#| label: Asia

Asia_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Asia") |>
  mutate(
    sex = if_else(
      str_detect(Series, "Male"), "Male", "Female")
    ) |>
  filter(Series == "Intentional homicide rates per 100,000")

Asia_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Asia"
  ) + theme_bw()
```

While homicide victims in Asia has decreased since the first year the data was recorded, it is evident that the declines have gotten steeper as the years have gone on.

```{r}
#| label: Europe

Europe_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Europe") |>
  mutate(
    sex = if_else(
      str_detect(Series, "Male"), "Male", "Female")
    ) |>
  filter(Series == "Intentional homicide rates per 100,000")

Europe_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Europe"
  ) + theme_bw()
```

Similar to Asia, Europe has also seen a decrease in their homicide rate; however, the rates of the decline have varied. From 2005 - 2010, there was a steep decline, which seemed to have leveled off from 2010-2015. From 2015 onward, the number of homicide victims seems to have decreased but at a much slower rate.

```{r}
#| label: Oceania

Oceania_homicides_and_crime_un <- homicides_and_crime_un |>
  filter(X == "Oceania") |>
   mutate(
    sex = if_else(
      str_detect(Series, "Male"), "Male", "Female")
    ) |>
  filter(Series == "Intentional homicide rates per 100,000")

Oceania_homicides_and_crime_un |>
  ggplot(aes(x = Year, y = Value, group = 1)) +
  geom_line() +
  geom_point() + 
  labs(
    x = "Year",
    y = "Homicide Victims Per 100k",
    title = "Intentional Homicide Victims in Oceania"
  ) + theme_bw()
```

Oceania saw a steep decline in the homicide rates from 2005-2010, which seemed to level off from 2010-2015, but the homicide victims have skyrocketed since 2015.

**TBD:**

Regional Regressions

Graph Modifications/Tidying up -- Make em' colorful and nice :)

**Results:**
